The Elements of Theses

### 2014-01-23 thesis/DataManagement.tex

\chapter{VM-centric Storage Management with Approximate Deletion}
\label{chap:data}
\section{Introduction}
\label{data:intro}
In this chapter, we discuss a low-cost architecture option that collocates
a backup service with other cloud services and uses a minimum amount of resources. 
We also consider the fact that after
deduplication, most data chunks are shared by several to many virtual machines.
Failure of a few shared data chunks can have a 
broad effect and many
snapshots of virtual machines could be affected.
The previous work in deduplication focuses on the efficiency and approximation of
fingerprint comparison, and has not addressed fault tolerance issues  together with deduplication.
Thus we also seek deduplication options that yield better fault isolation.
Another issue considered is that
that garbage collection after deletion of old snapshots also competes for computing resources. 
Sharing of data chunks among by multiple VMs needs to be detected during
garbage collection and such dependencies complicate deletion operations.

@@

\section{Background and Design Considerations}
\label{data:options}
Our key design consideration is VM dependence minimization during deduplication 
and file system block management.
\begin{itemize}
\item {\em Deduplication localization.}
Because a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, only one copy of duplicates is stored in the storage,
this artificially creates data dependencies among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen periodically 
in a large-scale cloud and
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can increase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
discouraged. 
Another disadvantage of sharing is that it complicates snapshot deletion, 
which  occurs frequently when snapshots expire regularly. 
The mark-and-sweep approach~\cite{Guo2011,Fabiano2013}  is effective for deletion, but still carries a significant cost
to count if a data chunk is still shared by other snapshots. 
Localizing deduplication can  minimize data sharing and simplify deletion while sacrificing 
deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\item{\em  Management of file system blocks.}
The file system block (FSB) size in a distributed file system such as  Hadoop and GFS is uniform and large (e.g.  64MB),
while the data chunk in a typical deduplication system is of a non-uniform size with 4KB or 8KB on average.
Packaging data chunks to an FSB can create more data dependencies among VMs
since a file system block can be shared by even more VMs.
Thus we need to consider a minimum association of FSBs to VMs in the packaging process.
\end{itemize}

@@

Three API calls are supported for data backup:
%\begin{itemize}

@@

%\item
\noindent\textbf{Get()}.
The fetch operation for the PDS data chunk is straightforward since each reference contains 
the file offset, and the size of a PDS chunk is available from a segment recipe.
We also maintain a small data cache for the PDS data service to speedup common data fetching.

To read a non-PDS chunk using its reference with container ID and local chunk ID,  the snapshot store client first loads the
corresponding VM's container index file specified by the container ID, then searches the chunk
groups  using their  chunk ID coverage.
After that, it reads the identified chunk group from DFS, decompresses it, and seeks to the exact chunk data 
specified by the chunk ID. 
Finally, the client updates its internal chunk data cache with the newly loaded content to 
anticipate future sequential reads.

@@

@@

\subsection{Impact on Deduplication Efficiency}
Choosing the value $k$ for the most popular chunks affects the deduplication efficiency.
We analyze this impact based on the characteristics  of the VM snapshot traces
studied from  application datasets.
A previous study shows that the popularity of data chunks after local deduplication follows 
a Zipf-like distribution\cite{Breslau1999a} and its
exponent $\alpha$ is ranged between 0.65  and  0.7~\cite{WeiZhangIEEE}. 
Figure~\ref{fig:zipf-data} illustrates the Zipf-like distribution of chunk popularity.

@@

\section{Related Work}
\label{data:related}
Since it is expensive to compare a large number of chunk signatures for deduplication,
several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR}.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements with the tradeoff of a reduced deduplication ratio.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault tolerance in the cluster-based environment that we have considered
in this work.

In designing a VC duplication solution, we have considered and adopted some of
the following previously-developed techniques.
1)
{\em Changed block Tracking}.
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011}.
Such a scheme  is  VM-centric since deduplication is localized. 
We are seeking for a tradeoff since 
global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
2) {\em Stateless Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing by Dong et al.~\cite{Dong2011} 
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index}.
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

>>>>

\chapter{VM-centric Storage Management with Approximate Deletion}
\label{chap:data}
\section{Introduction}
\label{data:intro}
In this chapter, we discuss a low-cost storage and management architecture that collocates
a backup service with other cloud services and uses a minimum amount of resources. 
We also consider the fact that after
deduplication, most data chunks are shared by several to many virtual machines.
Failure of a few shared data chunks can have a 
broad effect and many
snapshots of virtual machines could be affected.
The previous work in deduplication focuses on the efficiency and approximation of
fingerprint comparison, and has not addressed fault tolerance issues  together with deduplication.
Thus we also seek deduplication options that yield better fault isolation.
Another issue considered is that
that garbage collection after deletion of old snapshots also competes for computing resources. 
Sharing of data chunks among by multiple VMs needs to be detected during
garbage collection and such dependencies complicate deletion operations.

@@

\section{Design Considerations}
\label{data:options}
Our key design consideration is VM dependence minimization during deduplication 
and file system block management.
\begin{itemize}
\item {\em Deduplication localization.}
Because a data chunk is compared with fingerprints collected from all VMs during
the deduplication process, only one copy of duplicates is stored in the storage,
this artificially creates data dependencies among different VM users. 
Content sharing via deduplication affects fault isolation since machine failures happen periodically 
in a large-scale cloud and
loss of a small number of shared data chunks can 
cause the unavailability of snapshots for a large number of virtual machines.
Localizing the impact of deduplication can increase fault isolation and resilience.
Thus from the fault tolerance point of view,  duplicate sharing among multiple VMs is 
discouraged. 
Another disadvantage of sharing is that it complicates snapshot deletion, 
which  occurs frequently when snapshots expire regularly. 
The mark-and-sweep approach~\cite{Guo2011,Fabiano2013}  is effective for deletion, but still carries a significant cost
to count if a data chunk is still shared by other snapshots. 
Localizing deduplication can  minimize data sharing and simplify deletion while sacrificing 
deduplication efficiency, and  can facilitate parallel execution of snapshot operations.
\item{\em  Management of file system blocks.}
The file system block (FSB) size in a distributed file system such as  Hadoop and GFS is uniform and large (e.g.  64MB),
while the data chunk in a typical deduplication system is of a non-uniform size with 4KB or 8KB on average.
Packaging data chunks to an FSB can create more data dependencies among VMs
since a file system block can be shared by even more VMs.
Thus we need to consider a minimum association of FSBs to VMs in the packaging process.
\end{itemize}

@@

Our snapshot store supports three API calls for block data operations:
%\begin{itemize}

@@

%\item
\noindent\textbf{Get()}.
The fetch operation for the PDS data chunk is straightforward since each reference contains 
the file offset, and the size of a PDS chunk is available from a segment recipe.
We also maintain a small data cache for the PDS data service to speedup common data fetching.
To read a non-PDS chunk using its reference with container ID and local chunk ID,  the snapshot store client first loads the
corresponding VM's container index file specified by the container ID, then searches the chunk
groups  using their  chunk ID coverage.
After that, it reads the identified chunk group from DFS, decompresses it, and seeks to the exact chunk data 
specified by the chunk ID. 
Finally, the client updates its internal chunk data cache with the newly loaded content to 
anticipate future sequential reads.

@@

Additional APIs such as \textbf{Scan()}, \textbf{Compact()}, \textbf{Create()} and \textit{Remove()} are provided for 
higher level operations. Scan() is used by the map-reduce procedure to collect the most popular blocks among all snapshot stores.
Compact() is called when the system determines a container has too much unclaimed space and needs to be reclaimed.
Create() and Remove() initialize and delete a container respectively.

@@

\subsection{Impact on Deduplication Efficiency}
Choosing the value $k$ for the most popular chunks affects the deduplication efficiency.
We analyze this impact based on the characteristics  of the VM snapshot traces
studied from  application datasets.
A previous study shows that the popularity of data chunks after local deduplication follows 
a Zipf-like distribution\cite{Breslau1999a, Baek2011} and its
exponent $\alpha$ is ranged between 0.65  and  0.7~\cite{WeiZhangIEEE}. 
Figure~\ref{fig:zipf-data} illustrates the Zipf-like distribution of chunk popularity.

@@

\section{Related Work}
\label{data:related}
Since it is expensive to compare a large number of chunk signatures for deduplication,
several techniques have been proposed to speedup searching of duplicate
fingerprints. For example, the data domain method ~\cite{bottleneck08} 
uses  an in-memory Bloom filter and a prefetching cache for data chunks  which may be
accessed.  An improvement to this work with parallelization is in ~\cite{MAD210,DEBAR,arrow08}.
The approximation techniques are studied in~\cite{extreme_binning09,Guo2011,WeiZhangIEEE}  
to reduce memory requirements with the tradeoff of a reduced deduplication ratio.
Additional inline deduplication techniques are studied in ~\cite{sparseindex09,Guo2011,Srinivasan2012}
and a parallel batch solution for cluster-based deduplication is 
studied in ~\cite{wei2013}. 
All of the above approaches have focused on optimization of deduplication
efficiency, and none of them have considered the impact
of deduplication on fault tolerance in the cluster-based environment that we have considered
in this work.

In designing a VC duplication solution, we have considered and adopted some of
the following previously-developed techniques.
1)
{\em Changed block Tracking}.
VM snapshots can be  backed up  incrementally by identifying data segments that have
changed from the previous version of the snapshot~\cite{Clements2009,Vrable2009,TanIPDPS2011,cvfs03}.
Such a scheme  is  VM-centric since deduplication is localized. 
We are seeking for a tradeoff since 
global signature comparison can deliver additional compression~\cite{Guo2011,Dong2011,extreme_binning09}.
2) {\em Stateless Data Routing}.
One approach for scalable duplicate comparison is to use a content-based hash
partitioning algorithm called stateless data routing by Dong et al.~\cite{Dong2011} 
Stateless data routing divides the deduplication work with a similarity approximation. This work 
is similar to Extreme Binning by Bhagwat et al.~\cite{extreme_binning09} and 
each request is routed  to a machine which holds
a Bloom filter  or can fetch on-disk index for additional comparison.
While this approach is VM-oblivious, it motivates us to  use  a combined signature of a dataset to narrow
VM-specific local search.
3) {\em Sampled Index}.
One effective approach that reduces memory usage is 
to use a sampled index with prefetching, proposed  by Guo and Efstathopoulos\cite{Guo2011}. 
The algorithm is VM oblivious and it is not easy  to adopt for a distributed architecture. 
To use a distributed memory version of the sampled index, every deduplication request
may access a remote machine for index lookup and the overall overhead of access latency for all requests
can be significant.

### 2014-01-23 thesis/InlineDedup.tex

Our synchronous batched deduplication introduced in chapter~\ref{chap:offline} does not apply 
to such situation which requests fast inline deduplication.
Based on the requirement that we are now pursuing low-cost solution with short job turnaround time,
we choose to design a deduplication solution which will sacrifice the deduplication efficiency to 
satisfy the cost and time constraints.
By observations on the VM snapshot data from production cloud, we found snapshot data duplication 
can be easily classified into two categories: \emph{inner-VM} and \emph{cross-VM}. Inner-VM duplication
exists between VM's snapshots, because the majority of data are unchanged during each backup period. 
On the other hand, Cross-VM duplication is mainly due to widely-used software and libraries such as Linux and MySQL.
As the result, different VMs tend to backup large amount of highly similar data.

@@

\section{Requirements and Design Options}
\label{inline:options}
We discuss the characteristics and 
main requirements for VM snapshot backup in a cloud environment.
which are different from a traditional data backup. 
% That arises mainly in Alibaba's Aliyun cloud service.
%concerns
%Base on Aliyun's production environment, the snapshot backup job has to satisfy 
\begin{enumerate}
\item {\em Cost consciousness.}
There are tens of thousands of VMs running on a large-scale cluster. 
The amount of data is so huge such that backup cost must be controlled carefully.
On the other hand, the computing resources allocated for snapshot service is very limited
because VM performance has higher priority.  
At Aliyun, it is required that while CPU and disk usage should be small or modest during backup time,
the memory footprint of snapshot service should not exceed 500MB at each node.

@@

The inner VM deduplication contains two levels of duplicate detection efforts and the representation of
each snapshot is correspondingly designed as a two-level level index data structure in the form of a hierarchical
directed acyclic graph as shown in Figure~\ref{fig:snapshot}.
An image file is divided into a set of segments and each  segment contains hundreds of content blocks from the bottom level.
These blocks are of variable-size, partitioned using
the standard chunking technique~\cite{similar94} with 4KB as the average block size. 
%To simplify the deduplication process, segments are aligned to fix-sized boundaries, currently using 2MB.
Segment metadata (called segment recipe) records its  content block hashes and data pointers. 
The snapshot recipe contains a list of segments and other meta data information.
\begin{itemize}
\item {\em Level 1 Changed segment tracking}.
% If a segment is not changed, indicated by a dirty bit embedded in the virtual disk driver, its content blocks are not changed as well, thus its segment recipe can be simply reused. Operations at this level have almost no cost and most of unmodified data are filtered here. 
We start with the changed block tracking approach in a coarse grain segment level.
In our implementation with Xen on an Alibaba platform, the segment size is 2MB
and the device driver is extended to support tracking changed segments using a dirty bitmap. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement such changed block tracking feature.

@@

How can we choose the length of a segment?
Instead of using variables-sized segments, we take a simple approach
to let every segment being aligned to the page boundary of each virtual image file.
For Aliyun, each VM image file is represented as a virtual disk file format
(called \emph{vhd} at Xen) and we use a dirty bit to capture if a page (or segment) of a virtual disk file 
has been modified or not to ease the segment-level deduplication.
%This is effective because  
%a vhd   file keeps the allocated inner blocks at the same positions until deletion,
%which means the locality of snapshot data is almost natively aligned. 
A dirty bit array is used to indicate which segments have been modified or not. 
Each page (segment) size in our implementation uses 2MB, which contains a large number of content blocks.
%Enforcing a boundary at every 2MB will 
%only break 0.2\% of total data blocks which is tiny.
% dirty bits array. This is because
%Inside the TapDisk driver, we maintain an array of \emph{dirty bits} to record the dirty area
%of runtime VM image file since the last snapshot, each bit represent a 2MB fix-sized
%data segment. During a snapshot operation, we only need to look at the dirty region
%rather than scan over the whold image file, which would be extremely slow.

@@

Once level-2 deduplication is activated for those segments that have been modified,
it requires memory to load  block fingerprints from the corresponding
parent snapshot's segment.
This scheme processes one segment at time and each segment of 2MB contains about 
500 content blocks on average given 4KB is the average block size.
That only takes a tiny amount of space to hold their fingerprints.
%Given  1-2 hours of window to backup snapshots and every VM has  on average about 40GB, there is enough
%time to search all segments sequentially~\cite{NOT-CLEAR-HERE}

@@

The PDS collected from user-related data is generally proportional to the data size.
As discussed in Section~\ref{inline:impl}, selecting about 1\% of data (after level 1 and level 2 deduplication)
can cover about 38\% of data blocks. Consider we allow maximum 25 VMs per machine, each VM has
about 30GB of user-related data, having 10 snapshots in its snapshot store and the data change ratio during
each backup is 10\%, this translates to 15GB PDS data per machine. 
Consider each PDS hash index entry cost about 40 bytes and the average block size is 4KB, this leads to
a 1:100 ratio so the 
memory cost by PDS hash index is about 150MB per machine.
On the other hand, the PDS from OS-related data is only relevant to the number of OS releases.
Our experiment on 7 major OS releases shows that about 100GB of data is never modified, and we expect
it won't grow over 200GB in the near future. So it would cost the entire cluster 2GB of memory to
store its hash index, or 20MB per machine on a 100 node cluster. On average each OS disk has about 10GB
of data that can be completely eliminated in this way.
Thus in total the PDS index size per node takes less than 170MB in a large cluster bigger than 100 nodes, 
covering over 47\% of blocks in OS and data disks after inner-VM deduplication.
This memory usage is well below the limit required by our VM services.

@@

At Aliyun our target is to backup cluster up to 1000 nodes
with 25 VMs on each.
%We are running our deduplication/backup  service on 100 nodes.
%Memory usage is about 150MB space per node during backup and
%the CPU usage is very small during the experiments. 
Based on the data studied,  each VM has about  40GB of storage  data usage on average,
OS disk and data disk each takes about 50\% of storage space.
The backup of VM snapshots is completed within two hours every day,
and that translates to a backup throughput of 139GB per second, or 500TB per hour.
For each VM, the system keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.

@@

Since it's impossible to perform large scale analysis without affecting the VM performance,
we sampled two data sets from real user VMs to measure the effectiveness of our deduplication scheme.
Dataset1 is used study the detail impact of 3-level deduplication process,
it compose of 35 VMs from 7 popular OSes: 
Debian, Ubuntu, Redhat, CentOS, Win2003 32bit, win2003 64 bit and win2008 64 bit. For each OS, 
5 VMs are chosen, and every VM come with 10 full snapshots of it OS and data disk. 
The overall data size for this 700 full snapshots is 17.6 TB.

@@

Thus with about 1\% of blocks on data disks, PDS can cover about 38\% of total data blocks
appeared in all data snapshots. The corresponding PDS index uses no more than 150MB memory per machine,
which can be easily co-located with other cloud services.

@@

Based on the above analysis, we have selected a small set of most popular
OS blocks, which is about 100GB OS data and its corresponding PDS index takes about 1GB memory space in total.
They can cover sufficiently over 70\% of OS-related data.

@@

%We rank unique data blocks by their duplication count, and choose the hottest blocks as PDS. 
Figure \ref{fig:datacdssize} shows the relationship between PDS cache size and relative space saving ratio
compared to the full deduplication.  The unit of PDS size is gigabytes.
We define \emph{space saving ratio} as the space saving of PDS method divided by 
full deduplication space reduction. 
With a 100GB PDS data (namely 1GB PDS index) can still  accomplish about 75\% 
of what perfect deduplication can do.
%But this effect decreases when more data are added to PDS. 
%The lower bound of PDS space saving ratio is 50\%, which is very easy to accomplish.

@@

Data deduplication techniques can eliminate redundancy globally among different files from different users. 
Backup systems have been developed to use content hash (finger prints) to identify duplicate 
content~\cite{venti02,Rhea2008}.
%,NGmiddleware2011}. 
Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper} 
use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
%Chunking divides a data stream into variable length chunks, it has been used to conserve bandwidth\cite{lbfs01}, 
%search and index documents in large repositories\cite{bhag07}, scalable storage systems\cite{hydrastor09}, 
%store and transfer large directory trees efficiently and with high reliability\cite{jumbo07}.
As data grows to be big, fingerprint lookup in such schemes
becomes too slow to be scalable.
%and searching such disk
%However, unless some form of locality or similarity is exploited, inline, chunk-based deduplication, 
%when done at a large scale faces what has been termed the disk bottleneck problem: to facilitate fast chunk ID lookup, 
%a single index containing the chunk IDs of all the backed up chunks must be maintained. 
%As the backed up data grows, the index overflows the amount of RAM available and must be paged to disk. 
%Without locality, the index cannot be cached effectively, and it is popular for nearly 
%every index access to require a random disk access. This disk bottleneck severely limits deduplication throughput.
Several techniques have been proposed to speedup searching of duplicate 
content. For example,  
Zhu et al.~\cite{bottleneck08} tackle it 
by using an in-memory Bloom filter and prefetch groups of chunk IDs that are likely to be 
accessed together with high probability. It takes significant memory resource for filtering and caching.
NG et al.~\cite{ NGmiddleware2011}  use  
a related filtering technique for integrating deduplication in Linux  file system and the memory
consumed is up to 2GB for a single machine. That is still too big in our context discussed below. 
%Lillibridge et al.~\cite{sparseindex09} break list of chunks 
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is 
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments. 
%These are segments that share many chunk IDs with the incoming segment with high probability.
%Deepavali et al.~\cite{extreme_binning09}  use signature-based file similarity  and group similar files
%into the same physical location (bins) to deduplicate against each other.

>>>>

Our synchronous batched deduplication introduced in chapter~\ref{chap:offline} does not apply 
to such situation which requires fast inline deduplication.
Based on the requirement that we are now pursuing low-cost solution with short job turnaround time,
we choose to design a deduplication solution which will sacrifice the deduplication efficiency to 
satisfy the cost and time constraints.
By observations on the VM snapshot data from production cloud, we found snapshot data duplication 
can be easily classified into two categories: \emph{inner-VM} and \emph{cross-VM}. Inner-VM duplication
exists between VM's snapshots, because the majority of data are unchanged during each backup period. 
On the other hand, Cross-VM duplication is mainly due to widely-used software and libraries such as Linux and MySQL.
As the result, different VMs tend to backup large amount of highly similar data.

@@

\section{Requirements and Design Options}
\label{inline:options}
We discuss the characteristics and 
main requirements for VM snapshot backup in a cloud environment.
which are different from a traditional data backup. 
% That arises mainly in Alibaba's Aliyun cloud service.
%concerns
%Base on Aliyun's production environment, the snapshot backup job has to satisfy 
\begin{enumerate}
\item {\em Cost consciousness.}
There are tens of thousands of VMs running on a large-scale cluster. 
The amount of data is so huge such that backup cost must be controlled carefully.
On the other hand, the computing resources allocated for snapshot service is very limited
because VM performance has higher priority.  
At Aliyun, it is required that while CPU and disk usage should be small or modest during backup time,
the memory footprint of snapshot service should not exceed 500 MB at each node.

@@

The inner VM deduplication contains two levels of duplicate detection efforts and the representation of
each snapshot is correspondingly designed as a two-level level index data structure in the form of a hierarchical
directed acyclic graph as shown in Figure~\ref{fig:snapshot}.
An image file is divided into a set of segments and each  segment contains hundreds of content blocks from the bottom level.
These blocks are of variable-size, partitioned using
the standard chunking technique~\cite{similar94} with 4KB as the average block size. 
%To simplify the deduplication process, segments are aligned to fix-sized boundaries, currently using 2MB.
Segment metadata (called segment recipe) records its  content block hashes and data pointers. 
The snapshot recipe contains a list of segments and other meta data information.
\begin{itemize}
\item {\em Level 1 Changed segment tracking}.
% If a segment is not changed, indicated by a dirty bit embedded in the virtual disk driver, its content blocks are not changed as well, thus its segment recipe can be simply reused. Operations at this level have almost no cost and most of unmodified data are filtered here. 
We start with the changed block tracking approach in a coarse grain segment level.
In our implementation with Xen on an Alibaba platform, the segment size is 2 MB
and the device driver is extended to support tracking changed segments using a dirty bitmap. 
Since every write for a segment will touch a dirty bit, the device driver maintains dirty bits in memory
and cannot afford a small segment size.
It should be noted that dirty bit tracking is supported or can be easily implemented in 
major virtualization solution vendors. For example,
the VMWare hypervisor has an API to let external backup applications know 
the changed areas since last backup. 
The Microsoft SDK provides an API that allows external applications to monitor 
the VM's I/O traffic and implement such changed block tracking feature.

@@

How can we choose the length of a segment?
Instead of using variables-sized segments, we take a simple approach
to let every segment being aligned to the page boundary of each virtual image file.
For Aliyun, each VM image file is represented as a virtual disk file format
(called \emph{vhd} at Xen) and we use a dirty bit to capture if a page (or segment) of a virtual disk file 
has been modified or not to ease the segment-level deduplication.
%This is effective because  
%a vhd   file keeps the allocated inner blocks at the same positions until deletion,
%which means the locality of snapshot data is almost natively aligned. 
A dirty bit array is used to indicate which segments have been modified or not. 
Each page (segment) size in our implementation uses 2 MB, which contains a large number of content blocks.
%Enforcing a boundary at every 2MB will 
%only break 0.2\% of total data blocks which is tiny.
% dirty bits array. This is because
%Inside the TapDisk driver, we maintain an array of \emph{dirty bits} to record the dirty area
%of runtime VM image file since the last snapshot, each bit represent a 2MB fix-sized
%data segment. During a snapshot operation, we only need to look at the dirty region
%rather than scan over the whold image file, which would be extremely slow.

@@

Once level-2 deduplication is activated for those segments that have been modified,
it requires memory to load  block fingerprints from the corresponding
parent snapshot's segment.
This scheme processes one segment at time and each segment of 2 MB contains about 
500 content blocks on average given 4KB is the average block size.
That only takes a tiny amount of space to hold their fingerprints.
%Given  1-2 hours of window to backup snapshots and every VM has  on average about 40GB, there is enough
%time to search all segments sequentially~\cite{NOT-CLEAR-HERE}

@@

The PDS collected from user-related data is generally proportional to the data size.
As discussed in Section~\ref{inline:impl}, selecting about 1\% of data (after level 1 and level 2 deduplication)
can cover about 38\% of data blocks. Consider we allow maximum 25 VMs per machine, each VM has
about 30 GB of user-related data, having 10 snapshots in its snapshot store and the data change ratio during
each backup is 10\%, this translates to 15 GB PDS data per machine. 
Consider each PDS hash index entry cost about 40 bytes and the average block size is 4KB, this leads to
a 1:100 ratio so the 
memory cost by PDS hash index is about 150 MB per machine.
On the other hand, the PDS from OS-related data is only relevant to the number of OS releases.
Our experiment on 7 major OS releases shows that about 100 GB of data is never modified, and we expect
it won't grow over 200 GB in the near future. So it would cost the entire cluster 2 GB of memory to
store its hash index, or 20 MB per machine on a 100 node cluster. On average each OS disk has about 10 GB
of data that can be completely eliminated in this way.
Thus in total the PDS index size per node takes less than 170 MB in a large cluster bigger than 100 nodes, 
covering over 47\% of blocks in OS and data disks after inner-VM deduplication.
This memory usage is well below the limit required by our VM services.

@@

At Aliyun our target is to backup cluster up to 1000 nodes
with 25 VMs on each.
%We are running our deduplication/backup  service on 100 nodes.
%Memory usage is about 150MB space per node during backup and
%the CPU usage is very small during the experiments. 
Based on the data studied,  each VM has about  40 GB of storage  data usage on average,
OS disk and data disk each takes about 50\% of storage space.
The backup of VM snapshots is completed within two hours every day,
and that translates to a backup throughput of 139 GB per second, or 500TB per hour.
For each VM, the system keeps 10 automatically-backed snapshots in the storage while
a user may instruct extra snapshots to be saved.

@@

Since it's impossible to perform large scale analysis without affecting the VM performance,
we sampled two data sets from real user VMs to measure the effectiveness of our deduplication scheme.
Dataset1 is used study the detail impact of 3-level deduplication process,
it compose of 35 VMs from 7 popular OSes: 
Debian, Ubuntu, Redhat, CentOS, Win2003 32bit, Win2003 64 bit and Win2008 64 bit. For each OS, 
5 VMs are chosen, and every VM come with 10 full snapshots of it OS and data disk. 
The overall data size for this 700 full snapshots is 17.6 TB.

@@

Thus with about 1\% of blocks on data disks, PDS can cover about 38\% of total data blocks
appeared in all data snapshots. The corresponding PDS index uses no more than 150 MB memory per machine,
which can be easily co-located with other cloud services.

@@

Based on the above analysis, we have selected a small set of most popular
OS blocks, which is about 100 GB OS data and its corresponding PDS index takes about 1 GB memory space in total.
They can cover sufficiently over 70\% of OS-related data.

@@

%We rank unique data blocks by their duplication count, and choose the hottest blocks as PDS. 
Figure \ref{fig:datacdssize} shows the relationship between PDS cache size and relative space saving ratio
compared to the full deduplication.  The unit of PDS size is gigabytes.
We define \emph{space saving ratio} as the space saving of PDS method divided by 
full deduplication space reduction. 
With a 100 GB PDS data (namely 1 GB PDS index) can still  accomplish about 75\% 
of what perfect deduplication can do.
%But this effect decreases when more data are added to PDS. 
%The lower bound of PDS space saving ratio is 50\%, which is very easy to accomplish.

@@

Data deduplication techniques can eliminate redundancy globally among different files from different users. 
Backup systems have been developed to use content hash (finger prints) to identify duplicate 
content~\cite{venti02,Rhea2008}.
%,NGmiddleware2011}. 
Today's commercial data backup systems (e.g. from EMC and NetApp)
%\cite{emc_avamar}\cite{datadomain_whitepaper} 
use a variable-size chunking algorithm to detect duplicates in file data~\cite{similar94,hydrastor09}.
%Chunking divides a data stream into variable length chunks, it has been used to conserve bandwidth\cite{lbfs01}, 
%search and index documents in large repositories\cite{bhag07}, scalable storage systems\cite{hydrastor09}, 
%store and transfer large directory trees efficiently and with high reliability\cite{jumbo07}.
As data grows to be big, fingerprint lookup in such schemes
becomes too slow to be scalable.
%and searching such disk
%However, unless some form of locality or similarity is exploited, inline, chunk-based deduplication, 
%when done at a large scale faces what has been termed the disk bottleneck problem: to facilitate fast chunk ID lookup, 
%a single index containing the chunk IDs of all the backed up chunks must be maintained. 
%As the backed up data grows, the index overflows the amount of RAM available and must be paged to disk. 
%Without locality, the index cannot be cached effectively, and it is popular for nearly 
%every index access to require a random disk access. This disk bottleneck severely limits deduplication throughput.
Several techniques have been proposed to speedup searching of duplicate 
content. For example,  
Zhu et al.~\cite{bottleneck08} tackle it 
by using an in-memory Bloom filter and prefetch groups of chunk IDs that are likely to be 
accessed together with high probability. It takes significant memory resource for filtering and caching.
NG et al.~\cite{ NGmiddleware2011}  use  
a related filtering technique for integrating deduplication in Linux  file system and the memory
consumed is up to 2 GB for a single machine. That is still too big in our context discussed below. 
%Lillibridge et al.~\cite{sparseindex09} break list of chunks 
%into large segments, the chunk IDs in each incoming segment are sampled and the segment is 
%deduplicated by comparing with the chunk IDs of only a few carefully selected backed up segments. 
%These are segments that share many chunk IDs with the incoming segment with high probability.
%Deepavali et al.~\cite{extreme_binning09}  use signature-based file similarity  and group similar files
%into the same physical location (bins) to deduplicate against each other.

### 2014-01-23 thesis/Introduction.tex

The cost of supporting a large number of concurrent backup streams is high
because of the huge storage demand. 
If such VM snapshot data are plainly backuped without any duplicate detection, the storage waste would be 
extreme. For example, a fresh Windows server 2008 installation costs 25 GB on the virtual  disk,
and they are almost certianly duplicated with other Windows VM instances.
Using a separate  backup service with full deduplication support~\cite{venti02,bottleneck08}
can effectively identify and remove content duplicates among snapshots, 
but such a solution can be expensive. There is also a large amount of 
network traffic to transfer  data from the host machines to the backup facility
before duplicates are removed.

Unlike the previous work dealing with general file-level
backup and deduplication, our problem is focused on virtual
disk image backup. Although we treat each virtual disk as a
file logically, its size is very large. On the other hand, we need
to support parallel backup of a large number of virtual disks
in a cloud every day. One key requirement we face at busy clouds
 is that VM snapshot backup should only use a minimal
amount of system resources so that most of resources is kept
for regular cloud system services or applications. Thus our
objective is to exploit the characteristics of VM snapshot data
and pursue a cost-effective deduplication solution. Another
goal is to decentralize VM snapshot backup and localize
deduplication as much as possible, which brings the benefits
for increased parallelism and fault isolation.

@@

\textit{It  is  possible  to  build  an efficient, scalable, and fault tolerant backup storage
supporting highly accurate deduplication and easy management of deduplicated data
with sustainable high throughput.}

>>>>

The cost of supporting a large number of concurrent backup streams is high
because of the huge storage demand. 
If such VM snapshot data are plainly backed up without any duplicate reduction, storage waste would be 
extreme high. For example, a fresh Windows server 2008 installation costs 25 GB on the virtual disk,
and they are almost certianly duplicated with other Windows VM instances\cite{common07, pedal96}.
Using a separate  backup service with full deduplication support~\cite{venti02,bottleneck08}
can effectively identify and remove content duplicates among snapshots, 
but such a solution can be expensive. There is also a large amount of 
network traffic to transfer  data from the host machines to the backup facility
before duplicates are removed.

Unlike the previous work dealing with general file-level
backup and deduplication inside a centralized storage appliance, our problem is focused on virtual
disk image backup in a cluster of machines. Although we treat each virtual disk as a
file logically, its size is very large. On the other hand, we need
to support parallel backup of a large number of virtual disks
in a cloud every day. One key requirement we face at busy clouds
 is that VM snapshot backup should only use a minimal
amount of system resources so that most of resources is kept
for regular cloud system services or applications. Thus our
objective is to exploit the characteristics of VM snapshot data
and pursue a cost-effective deduplication solution. Another
goal is to decentralize VM snapshot backup and localize
deduplication as much as possible, which brings the benefits
for increased parallelism and fault isolation.

@@

\textit{It  is  possible  to  build  an efficient, scalable, and fault tolerant backup storage
supporting highly accurate deduplication with sustainable high throughput
and easy management on deduplicated data.}
